---
title: "Halloween Candy Project"
author: "Neil Mc"
date: "15/11/2019"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Halloween Candy Analysis


- Review of steps taken
- What I did not complete
- Step through Code
- Talk through any issues

## Steps taken

- read in 3 candy files
- for each file perform the following
  - select only those columns we require for analysis
  - pivot longer on candy (creating two columns candy and response)
  - rename columns to standard
  - add required columns (2015 file did not have all columns)
  - rearrange order of columns so each of 3 files is idenical
  
- bind the three long files into one (915,000 rows) 
- add a column "score". set to 1, -1 or 0 depending on response
```{r}
library(tidyverse)
library(dplyr)
library(readxl)
library(here)
library(janitor)
```

## Read the files
```{r read, echo=TRUE}
candy_data_raw_2015 <- 
        read_xlsx(here("raw_data/boing-boing-candy-2015.xlsx"))
candy_data_raw_2016 <- 
        read_xlsx(here("raw_data/boing-boing-candy-2016.xlsx"))
candy_data_raw_2017 <- 
        read_xlsx(here("raw_data/boing-boing-candy-2017.xlsx"))
dim(candy_data_raw_2015)
dim(candy_data_raw_2016)
dim(candy_data_raw_2017)
```

